---
title: "Lobster Exploration"
author: "Desik Somasundaram"
date: "8/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
```

## Read in the Lobster Abundance Data

```{r}
lobsters <- read_csv(here("data","Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>% 
  clean_names() %>% 
  uncount(count)
```

The uncount function essentially turns single rows with frequency counts into multiple row so that each row only represents one observation. It also gets rid of the count column.

Summary table that finds the total counts and mean carapace lengths of lobsters observed in the dataset by the site and year

```{r}
summary_lobs <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(mean_length = mean(size_mm, na.rm = TRUE), n = n())

ggplot(summary_lobs, aes(x = year, y = n)) + geom_point() + facet_wrap(~site)

```

## Proportion of legal lobsters at each site for 2020

```{r}
lobsters_2020 <- lobsters %>% 
  filter(year == "2020") %>% 
  mutate(legal = case_when(size_mm > 79.76 ~ "yes", size_mm < 79.76 ~ "no")) %>% 
  group_by(site, legal) %>% 
  summarize(count = n())

ggplot(lobsters_2020, aes(x = site, y = count, fill = factor(legal))) + geom_col(position = "fill")

```

## Random Lobster Wrangling

```{r}
ex_a <- lobsters %>% 
  filter(!site %in% c("AQUE", "MOHK"))

unique(ex_a$site)
```
```{r}
ex_b <- lobsters %>% 
  filter(month == 8)

unique(ex_b$month)
```
```{r}
ex_c <- lobsters %>% 
  filter(site == "AQUE" | size_mm > 70)
```

```{r}
ex_d <- lobsters %>% 
  filter(!site == "NAPL")

unique(ex_d$site)
```
```{r}
ex_e <- lobsters %>% 
  group_by(site) %>% 
  summarize(mean_length = mean(size_mm, na.rm = TRUE), sd_length = sd(size_mm, na.rm = TRUE))
```


```{r}
ex_f <- lobsters %>% 
  group_by(site, month) %>% 
  summarize(max_length = max(size_mm, na.rm = TRUE))
```

```{r}
ex_g <- lobsters %>% 
  mutate(size_cm = size_mm/100)
```

```{r}
ex_h <- lobsters %>% 
  mutate(site = tolower(site))
```

```{r}
ex_i <- lobsters %>% 
  mutate(area = as.character(area))

class(ex_i$area)

```
```{r}
ex_j <- lobsters %>% 
  mutate(size_bin = case_when(size_mm <= 70 ~"small", size_mm > 70 ~"large"))
```

```{r}
ex_k <- lobsters %>% 
  mutate(designation = case_when(site == "IVEE"|site == "NAPL" ~"MPA", TRUE ~"not MPA"))
```

